---
title: "P8105 Hw1-hy2951"
author: "Haochen Yu"
output: github_document
date: "2025-09-20"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
#Problem 1
#Load the necessary packages
library(moderndive) #provide the data set
library(ggplot2) #plotting
library(scales) #for date_labels (switch the date from Chinese to English)

#Load the data set into the session
data("early_january_weather")

#Compute the values for inline reporting
n_rows <- nrow(early_january_weather) #number of observations
n_cols <- ncol(early_january_weather) #number of variables
mean_temp <- mean(early_january_weather$temp, na.rm = TRUE) #mean temperature without missing value

#Build a scatterplot
p_temp <- ggplot(early_january_weather,
                 aes(x = time_hour, y = temp, color = humid)) + #map the variables to axes and color
          geom_point(alpha = 0.7) + #reduce overplotting
          scale_x_datetime(labels = label_date(format = "%b %d", locale = "en")) +   #setup date label from Chinese to English
          labs(title = "NYC Early January Weather: Temperature vs Time",
               x = "Time (hourly)",
               y = "Temperature (째F)",
               color = "Humidity") # Add title and axis labels

p_temp #print the plot

ggsave("figs/temp_vs_time_humid.png", plot = p_temp, width = 8, height = 5, dpi = 300) #save the plot

```

Data set description
This data set contains hourly weather observations from NYC's Central Park station during early January. It has `r n_rows` rows and `r n_cols` columns. The variables are: `r toString(names(early_january_weather))`. The key variables are `time_hour`, `temp`, `humid`. The mean temperature over this period is `r round(mean_temp,1)` 째F

Interpretation of the figure:
The plot shows temperatures follow a daily cycle, being lower at night and higher during the day. Temperatures vary from below 30 째F to above 50 째F. The higher humidity is generally associated with lower temperatures. And lower humidity is generally associated with higher temperatures.







